---
title: 'Assignment #2'
author: "J. Hamski"
date: "February 11, 2016"
output: html_document
---

1. Page 69, #12
*A company with a fleet of trucks faces increasing maintenance costs as the age and milage of the trucks increase.*




2.

$y \propto x^3$

```{r}
y <- c(0,1,2,6,14,24,37,58,82,114)
x <- 1:10
y.model <- x^3

modelset <- as.data.frame(cbind(x, y, y.model))

library(ggplot2)

ggplot(modelset, aes(x)) + 
  geom_line(aes(y = y.model, colour = "Stated Proportion")) + 
  geom_line(aes(y = y, colour = "Actual Data"))
```
 
The dataset does support the stated proportionality model. It looks better as $y \propto x^3$

```{r}
y <- c(0,1,2,6,14,24,37,58,82,114)
x <- 1:10
y.model <- x^2

modelset.2 <- as.data.frame(cbind(x, y, y.model))

ggplot(modelset.2, aes(x)) + 
  geom_line(aes(y = y.model, colour = "Cubic Proportionality")) + 
  geom_line(aes(y = y, colour = "Actual Data"))
```

3.

*Lumber Cutters* Using diameter of tree in inches at waist height, develop a model that predicts board feet of lumber produced by that tree as a function of diameter in inches. 

```{r}
#radius of a pine tree in inches
x <- c(17, 19, 20, 23, 25, 28, 32, 38, 39, 41)/2

#board feet divided by 10
y <- c( 19, 25, 32, 57, 71, 113, 123, 252, 259, 294) * 10
```

Consider two separate assumptions allowing each to lead to a model. 

i. Assume that all trees are right-circular cylinders and are approximately the same height. 

When creating lumbar boards from a tree, you can increase the foot-length produced two ways - by having longer trees and having wider trees that allow more rectangular boards to fit within a circular cross-section of the tree. Therefore, we can hypothesize that board-feet is proportional to volume: $V = \pi r^2 h$.  

In this case, we assume that all the trees are the same height and set h = 1. The resulting model is $V = k\pi(r)^2 = k(r)^2$.

According to Wikipedia a board foot of lumber has a volume of $144 in^3$.

```{r}
y.vol <- y*144
r.sqrd <- x**2

k <- y.vol[10] / r.sqrd[10]


```




ii. Assume that all trees are right-circular cylinders and that the height of the tree is proportional to the diameter. 

```{r}



```



4.
*Discuss several factors we completely ignored in our anlaysis of the gasoline milage problem.*


5. 

**
