---
title: "Assignment 14"
author: "J. Hamski"
date: "May 15, 2016"
output: html_document
---

http://www.sce.carleton.ca/faculty/chinneck/po/Chapter12.pdf

http://optlab-server.sce.carleton.ca/POAnimations2007/BranchAndBound.html


A company is assembling a team to carry out a series of operations.
There are four members of the team: A, B, C and D, and four operations
to be carried out. Each team member can carry out exactly one
operation. All four operations must be carried out successfully for the
overall project to succeed, however the probability of a particular team
member succeeding in a particular operation varies, as shown in the
table below. For example, if the team members were assigned to
operations in the order ABCD, then the overall probability of successful
completion of the project is (0.9)(0.6)(0.85)(0.7) = 0.3213. If there is any
possible way that the team can be arranged such that the overall
probability of success exceeds 45%, then the manager will approve the
project. Will the manager approve the project? If yes, what is the
arrangement of the team that gives the highest probability of success?

```{r}

A <- c(0.9, 0.8, 0.9, 0.85)
B <- c(0.7, 0.6, 0.8, 0.7)
C <- c(0.85, 0.7, 0.85, 0.8)
D <- c(0.75, 0.7, 0.75, 0.7)

tasks <- rbind(A, B, C, D)
colnames(tasks) <- c("1", "2", "3", "4")
```

```{r}
root <- prod(apply(tasks, 2, function(x) max(x)))

get.values <- function(input){
  prod(tasks[input[1],1],
       tasks[input[2],2],
       tasks[input[3],3],
       tasks[input[4],4])}
```

```{r}

```

